---
title: "Analysis on First Two Folders"
author: Aidan Fauth
format: 
  html:
    embed-resources: true
---

## Set Up

```{r}
rm(list = ls())
library(tidyverse)
library(here)
library(lubridate)
library(hms)
```

### Data

```{r}
A001SD001 <- read_csv(here("output_A001_SD001.csv"), guess_max = 10000)
A002SD013 <- read_csv(here("output_A002_SD013.csv"), guess_max = 10000)
```

## A001_SD001

```{r}
A001SD001_long <- A001SD001 |> relocate(file) |>
  pivot_longer(4:ncol(A001SD001), names_to = "species", values_to = "probability") |>
  mutate(date = str_sub(file, start = 1, end = 8),
          hour = str_sub(file, 10, 11),
          minute = str_sub(file, 12, 13),
          second = str_sub(file, 14, 15)) |>
  unite("date_time", c("date", "hour", "minute", "second"), sep = ":") |>
  mutate(date_time = ymd_hms(date_time),
          time = as_hms(date_time))

A001SD001_sightings <- A001SD001_long |> filter(!is.na(probability))

A001SD001_most_prob <- A001SD001_sightings |> group_by(file, start, end) |>
  ## filter(probability == max(probability)) |>
  filter(probability >= 0.75) |> 
  ungroup() |> ## minimum probability to show in plot
  mutate(species = fct_infreq(species)) |>
  mutate(species = fct_rev(species))

A001SD001_most_prob

ggplot(data = A001SD001_most_prob, aes(x = time, y = species)) +
  geom_jitter(aes(colour = probability), alpha = 0.5, height = 0.25) +
  theme_minimal() +
  scale_colour_viridis_c()
```

#### Facet by day:

Need to parse date_time variable first to get just the date (not the time)

```{r}
A001SD001_most_prob <- A001SD001_most_prob |> mutate(date = 
    format(date_time, "%m-%d"))
```

Now let's facet by day

```{r}
#| eval: false
ggplot(data = A001SD001_most_prob, aes(x = time, y = species)) +
  geom_jitter(aes(colour = probability), alpha = 0.5, height = 0.25) +
  facet_wrap(~date)+
  theme_minimal() +
  scale_colour_viridis_c()
```

That's way too crazy... let's look at a subset of the days and get rid of some species (like the top 10)?

```{r}
A001SD001_most_prob |> group_by(species) |> 
    summarise(n_recordings = n()) |> 
    arrange(desc(n_recordings)) |> 
    slice(1:10)
```

Save top 10 species as a new df

NOTE: some of these species are not native

```{r}
A001SD001_top_10 <- A001SD001_most_prob |>
     filter(species == "Pseudacris crucifer_Spring Peeper" |
species == "Aix sponsa_Wood Duck" |
species == "Cacomantis variolosus_Brush Cuckoo" |
species == "Catharus guttatus_Hermit Thrush" | 
species == "Sayornis phoebe_Eastern Phoebe" |
species == "Turdus migratorius_American Robin" | 
species == "Turdus philomelos_Song Thrush" |
species == "Manorina melanophrys_Bell Miner" | 
species == "Tapera naevia_Striped Cuckoo" | 
species == "Palmeria dolei_Akohekohe")

unique(A001SD001_top_10$species)
```

Now Let's grab a subset of the days (let's grap 3 on either side of the eclipse)

```{r}
A001SD001_plot_7d <- A001SD001_top_10 |> filter(date == "04-05" |
    date == "04-06" |
    date == "04-07" |
    date == "04-08" |
    date == "04-09" |
    date == "04-10" |
    date == "04-11")
```

Now let's make the same faceting plot but with the filtered df

```{r}
ggplot(data = A001SD001_plot_7d, aes(x = time, y = species)) +
  geom_jitter(aes(colour = probability), alpha = 0.5, height = 0.25) +
  facet_wrap(~date)+
  theme_minimal() +
  scale_colour_viridis_c()
```

Maybe grab two on either side

```{r}
A001SD001_plot_5d <- A001SD001_top_10 |> filter(date == "04-06" |
    date == "04-07" |
    date == "04-08" |
    date == "04-09" |
    date == "04-10")
```

Plot

```{r}
ggplot(data = A001SD001_plot_5d, aes(x = time, y = species)) +
  geom_jitter(aes(colour = probability), alpha = 0.5, height = 0.25) +
  facet_wrap(~date)+
  theme_minimal() +
  scale_colour_viridis_c()
```

Nothing is really popping out.

Possibly look at the number of IDs over time?

Let's look just at the spring peeper for now

```{r}
speeper_IDs <- A001SD001_plot_5d |> filter(species == "Pseudacris crucifer_Spring Peeper") |> group_by(date_time) |> mutate(tot_IDs = n())
```

```{r}
ggplot(data = speeper_IDs, aes(x = time, y = tot_IDs))+
    geom_line()+
    facet_wrap(~date)
    theme_minimal()
```

Hard to interpret.

Maybe just within the day of the eclipse?

Take the top ten species over that day:

```{r}
# filter the top10 df to only include the day of the eclipse and have a count for each species
A001SD001_eday_10 <- A001SD001_top_10 |> 
    filter(date == "04-08") |> 
    group_by(time, species) |> 
    mutate(tot_IDs = n())

# Line plot faceted by species
ggplot(data = A001SD001_eday_10, aes(x = time, y = tot_IDs)) +
    geom_line()+
    facet_wrap(~species)
    theme_minimal()
```


## A002_SD013

```{r}
A002SD013_long <- A002SD013 |> relocate(file) |>
  pivot_longer(4:ncol(A002SD013), names_to = "species", values_to = "probability") |>
  mutate(date = str_sub(file, start = 1, end = 8),
          hour = str_sub(file, 10, 11),
          minute = str_sub(file, 12, 13),
          second = str_sub(file, 14, 15)) |>
  unite("date_time", c("date", "hour", "minute", "second"), sep = ":") |>
  mutate(date_time = ymd_hms(date_time),
          time = as_hms(date_time))

A002SD013_sightings <- A002SD013_long |> filter(!is.na(probability))

A002SD013_most_prob <- A002SD013_sightings |> group_by(file, start, end) |>
  ## filter(probability == max(probability)) |>
  filter(probability >= 0.75) |> 
  ungroup() |> ## minimum probability to show in plot
  mutate(species = fct_infreq(species)) |>
  mutate(species = fct_rev(species))


ggplot(data = A002SD013_most_prob, aes(x = time, y = species)) +
  geom_jitter(aes(colour = probability), alpha = 0.5, height = 0.25) +
  theme_minimal() +
  scale_colour_viridis_c()
```

#### Facet by day:

Need to parse date_time variable first to get just the date (not the time)

```{r}
A002SD013_most_prob <- A002SD013_most_prob |> mutate(date = 
    format(date_time, "%m-%d"))
```

Get the top 10 species

```{r}
A002SD013_most_prob |> group_by(species) |> 
    summarise(n_recordings = n()) |> 
    arrange(desc(n_recordings)) |> 
    slice(1:10)
```

```{r}
A002SD013_top_10 <- A002SD013_most_prob |>
     filter(species == "Certhia americana_Brown Creeper" |
species == "Turdus migratorius_American Robin" |
species == "Sayornis phoebe_Eastern Phoebe" |
species == "Branta canadensis_Canada Goose" | 
species == "Aix sponsa_Wood Duck" |
species == "Poecile atricapillus_Black-capped Chickadee" | 
species == "Megaceryle alcyon_Belted Kingfisher" |
species == "Dryocopus pileatus_Pileated Woodpecker" | 
species == "Dryobates pubescens_Downy Woodpecker" | 
species == "Euphagus carolinus_Rusty Blackbird")
```

```{r}
A002SD013_plot_5d <- A002SD013_top_10 |> filter(date == "04-06" |
    date == "04-07" |
    date == "04-08" |
    date == "04-09" |
    date == "04-10")
```

Plot

```{r}
ggplot(data = A002SD013_plot_5d, aes(x = time, y = species)) +
  geom_jitter(aes(colour = probability), alpha = 0.5, height = 0.25) +
  facet_wrap(~date)+
  theme_minimal() +
  scale_colour_viridis_c()
```

Number of Species IDs on the day of the eclipse (top 10 species)

```{r}
# filter the top10 df to only include the day of the eclipse and have a count for each species
A002SD013_eday_10 <- A002SD013_top_10 |> 
    filter(date == "04-08") |> 
    group_by(time, species) |> 
    mutate(tot_IDs = n())

# Line plot faceted by species
ggplot(data = A002SD013_eday_10, aes(x = time, y = tot_IDs)) +
    geom_line()+
    facet_wrap(~species)
    theme_minimal()
```





