---
title: "Compiling a Main Data Frame"
format: html
---

## Goal:

In the following code, the output files from all the audio moth folders are compiled into a single 'main' data frame.


## Set up

```{r}
rm(list = ls())
library(tidyverse)
library(hms)
library(here)
```


## Compiling and Formatting Data Frame

Here we will take the folder containing all the output files and load them into list:

```{r}
file_list <- list.files(path = here("Output"), pattern = "no_list.csv$",)

# read in all the files as dfs
df_list <- map(file_list, \(x) read_csv(here("Output", x)))

# process each df to then bind_rows
cleaned_list <- map(df_list, \(x) x |> relocate(file) |>
  pivot_longer(4:ncol(x), names_to = "species", values_to = "probability") |>
  filter(!is.na(probability)))

# add a column with the AM name
for (i in 1:length(cleaned_list)){
  cleaned_list[[i]] <- cleaned_list[[i]] |> 
    mutate(id = str_sub(file_list[[i]], start = 8, end = 17))
}

# compile into single df
temp_df <- bind_rows(cleaned_list)
```


### Create date and time variables

```{r}
main_df <- temp_df |> mutate(date = str_sub(file, start = 1, end = 8),
          hour = str_sub(file, 10, 11),
          minute = str_sub(file, 12, 13),
          second = str_sub(file, 14, 15)) |>
  unite("date_time", c("date", "hour", "minute", "second"), sep = ":") |>
  mutate(date_time = ymd_hms(date_time),
          time = as_hms(date_time))
```


## Unite with Audio Moth Deployments

Take a look at the deployment df
```{r}
deploy <- read_csv(here("eclipse_audio_moth_deployments.csv"))
glimpse(deploy)
```

To merge the two data frames I need to create a new variable in the output df that will match with the `recorder_lab_id_num`

```{r}
main_df <- main_df |> mutate(AM_id = 
  as.numeric(str_extract(id, "[1-9][0-9]*")))
```

Now we can merge using a left join:

```{r}
merged_df <- left_join(main_df, deploy, 
join_by(AM_id == recorder_lab_id_num))
```

## Visualizations

Let's see the most common species across the entire array of AMs.

```{r}
merged_df |> group_by(species) |> 
  summarise(count = n()) |> 
  arrange(desc(count)) |> 
  slice(1:10)
```

Let's make a plot of the top 4 species:

```{r}
plot_df <- merged_df |> 
  filter(species == "Pseudacris crucifer_Spring Peeper" |
  species == "Branta canadensis_Canada Goose" |
  species == "Lithobates sylvaticus_Wood Frog" |
  species == "Lithobates palustris_Pickerel Frog")

ggplot(data = plot_df, aes(x = time, y = id)) +
  geom_jitter(aes(colour = probability), alpha = 0.5, height = 0.25) +
  facet_wrap(~species)+
  theme_minimal() +
  scale_colour_viridis_c()
```

Let's filter for the day of the eclipse:

```{r}
eclipse_day <- plot_df |> filter(as_date(date_time) == "2024-04-08")

ggplot(data = eclipse_day, aes(x = time, y = id)) +
  geom_jitter(aes(colour = probability), alpha = 0.5, height = 0.25) +
  facet_wrap(~species)+
  theme_minimal() +
  scale_colour_viridis_c()
```
