---
title: "Writing Species List"
author: "Aidan Fauth"
format: html
---

## Set up

```{r}
rm(list = ls())
library(tidyverse)
library(here)
```

## Data

```{r}
unfiltered_species <- read_csv(here("Species_lists/Ebird_species.csv"))
common_names <- read_csv("Species_lists/Species_verification.csv")
```

This code is to solve the problem that my species list only contains common names, but I have an unfiltered list from Ebird that has all scientific names as well.

This code will filter the rows of the 'unfiltered list' so that I have a csv that has both the common name and scientfic name.

Create an indicator column that matches the cleaned species list with the common names to the unfiltered list with both the common names and scientific names.

```{r}
species_ind <- unfiltered_species |> mutate(ind = if_else(Common_Name %in% common_names$Species, true = 1, false = 0)) |> 
    filter(ind == 1)
```

Now I can remove the indicator column and write the cleaned csv

```{r}
cleaned_species <- species_ind |> select(-ind)
```


## ERROR!

The scientific name has to be first, so I need to use separate and unite to fix this:

```{r}
cleaned_species <- cleaned_species |> separate(col = Combined_Name, 
into = c("C_name", "Sci_name"), sep = "_") |> 
    unite(col = "Combined_Name", c("Sci_name", "C_name"), sep = "_")    
```


Now let's check against the birdnet list

```{r}
birdnet <- read_csv(here("Species_lists/birdnet_species.csv"))

cleaned_species |> mutate(ind = if_else(Combined_Name %in% birdnet$Species, true = 1, false = 0)) |> 
    filter(ind == 0)
```

## Write the final csv

```{r}
write_csv(cleaned_species, here("Species_lists/cleaned_species.csv"))
```

## MORE PROBLEMS!!!

I continue to have problems with the species list but I learned about an ebird R package.

I'm going to try to ensure that the the species I have are the correct ones.

```{r}
library(auk)

species_file <- here("Species_lists/scientific_names_list.txt")

species_list <- readLines(species_file)

ebird_df <- ebird_species(species_list, type = "all")
```

Now I have a data frame of all my species and if they are in ebird taxonomy

Let's check if I have any NAs

```{r}
ebird_comb <- ebird_df |> select(2:3) |> unite(col = "Combined_name", sep = "_")

write_csv(ebird_comb, here("Species_lists/ebird_species_list.csv"))
```

## Anti-join with BirdNet Generated List

```{r}
cleaned <- read_csv(here("Species_lists/cleaned_species.csv"))
generated <- read_csv(here("Species_lists/generated_species.csv"))

```

Anti join:



