---
title: "Data Exploration"
format: html
---

## Goal

This file explores the birdnet output with the deployment data.

## Set up

```{r}
rm(list = ls())
library(tidyverse)
library(here)
library(hms)
merged_df <- read_csv(here("Data/merged_output_deployment.csv"))
glimpse(merged_df)
```

## Fixing datetime and time variable:

```{r}
merged_df <- merged_df |> mutate(date_time = date_time + seconds(start),
time = as_hms(time + start))
```

## Visualizations

Let's see the most common species across the entire array of AMs.

```{r}
merged_df |> group_by(species) |> 
  summarise(count = n()) |> 
  arrange(desc(count)) |> 
  slice(1:10)
```

Let's make a plot of the top 4 species:

```{r}
plot_df <- merged_df |> 
  filter(species == "Pseudacris crucifer_Spring Peeper" |
  species == "Branta canadensis_Canada Goose" |
  species == "Lithobates sylvaticus_Wood Frog" |
  species == "Turdus migratorius_American Robin")

ggplot(data = plot_df, aes(x = time, y = id)) +
  geom_jitter(aes(colour = probability), alpha = 0.5, height = 0.25) +
  facet_wrap(~species)+
  theme_minimal() +
  scale_colour_viridis_c()
```

Let's filter for the day of the eclipse:

```{r}
eclipse_day <- plot_df |> filter(as_date(date_time) == "2024-04-08")

ggplot(data = eclipse_day, aes(x = time, y = id)) +
  geom_jitter(aes(colour = probability), alpha = 0.5, height = 0.25) +
  facet_wrap(~species)+
  theme_minimal() +
  scale_colour_viridis_c()
```

Is there more of a trend with the higher probability identifications?

```{r}
eclipse_day_mp <- eclipse_day |> filter(probability > 0.75)

ggplot(data = eclipse_day_mp, aes(x = time, y = id)) +
  geom_jitter(aes(colour = probability), alpha = 0.5, height = 0.25) +
  facet_wrap(~species)+
  theme_minimal() +
  scale_colour_viridis_c()
```

### Focusing on eclipse time

We want select for the time-frame of the eclipse.

First, I want to look at the top 10 species that actually called during the eclipse day

```{r}
merged_df |> filter(as_date(date_time) == "2024-04-08") |> 
  group_by(species) |> 
  summarise(count = n()) |> 
  arrange(desc(count)) |> 
  slice(1:10)
```

Now instead of faceting by species, we will filter for one species and facet by day:

Let's get a general df for these graphs

MH Start

```{r}
days <- c(1:15)

## units are all in seconds since midnight

## create a vector of seconds that I want to add and subtract to and from
## the time of totality
interval_range <- (c(seq(30, 3030, by = 60)))
mid_point <- as_hms("15:25:30") |> as.numeric() ## time of totality

## convert time to seconds since midnight
merged_df <- merged_df |> mutate(time_basic = as_hms(time) |> as.numeric()) |>
  relocate(time_basic)
## only look at April 1 through 15
focused_time <- merged_df |> filter(day(date_time) %in% days)
```

```{r}
## filter hits (of any species) based on a provided "interval" value
## interval could be named better: it is really just the number of seconds
## we are adding and subtracting from the time of totality
## to create an interval
filter_interval <- function(mid_point, df, interval, min_prob) {
  df |> filter(time_basic >= mid_point - interval & time_basic <= mid_point + interval) |>
    mutate(interval_size = interval * 2) |>
    filter(probability >= min_prob)
}

## map through the function to get one data set filtered by each of the values in
## the interval_range variable

## goal: see how the number of hits changes as we get a wider and wider interval around the eclipse
## map through the function for the different interval sizes
dfs <- map_dfr(interval_range, filter_interval,
               mid_point = as_hms("15:25:30") |> as.numeric(),
               df = focused_time, min_prob = 0.75) |>
                   mutate(day = day(date_time),
                          month = month(date_time),
                          eclipse_ind = if_else(day == "8" & month == "4",
                                                true = "eclipse",
                                        false = "non-eclipse"))

## grab total number of detections (of any species) for each interval for
## each day for each month (only looking at April here, but wanted
## this to generalize to March if we included it).
df_count <- dfs |> group_by(day, month, interval_size) |>
  summarise(n_hits = n(),
    eclipse_ind = first(eclipse_ind))

## n_hits is cumulative for larger and larger interval sizes
## new_hits is the number of unique hits for each increasing interval 
## e.g., if there were 50 hits in the minute surrounding the eclipse and 
## 150 hits in the two minutes surrounding the eclipse, then
## n_hits would be (50, 200) and new_hits would be (50, 150)
df_count <- df_count |> group_by(day, month) |>
  mutate(n_hits_lag = lag(n_hits),
         n_hits_lag = if_else(is.na(n_hits_lag), true = 0, false = n_hits_lag)) |>
  mutate(new_hits = n_hits - n_hits_lag)

## this first one is rate
ggplot(data = df_count, aes(x = interval_size, y = n_hits)) +
  geom_line(aes(group = day, colour = eclipse_ind), alpha = 0.5) +
  scale_colour_viridis_d() +
  theme_minimal()

## this second one is just the raw number of new hits as the interval size
## surrounding the eclipse slowly increases by 1 minute on each side,
## centered at totality

## this suggests that there is an increase in overall call rate
## during the eclipse but that we are looking too far out
## the increasing amount of calls peters off as we get further from 
## totality and is no different than other days past
## an interval size of about 2000 seconds (which roughly 
## corresponds to adding and subtracting 30 minutes from the time of totality)
ggplot(data = df_count, aes(x = interval_size, y = new_hits)) +
  geom_line(aes(group = day, colour = eclipse_ind), alpha = 0.5) +
  scale_colour_viridis_d() +
  theme_minimal()

```

MH End

#### Spring Peeper

```{r}
sp <- focused_time |> filter(species == "Pseudacris crucifer_Spring Peeper") |> 
  filter(probability >= 0.75) # there are a lot of s. peeper points

ggplot(data = sp, aes(x = time, y = id)) +
  geom_jitter(aes(colour = probability), alpha = 0.5, height = 0.25) +
  facet_wrap(~day(date_time))+
  theme_minimal() +
  scale_colour_viridis_c()
```

For the rest let's try birds as they might be influenced differently

#### Robin

```{r}
robin <- focused_time |> filter(species == "Turdus migratorius_American Robin")

ggplot(data = robin, aes(x = time, y = id)) +
  geom_jitter(aes(colour = probability), alpha = 0.5, height = 0.25) +
  facet_wrap(~day(date_time))+
  theme_minimal() +
  scale_colour_viridis_c()
```

```{r}
focused_time |> ##filter(species == "Melospiza melodia_Song Sparrow") |> 
  filter(probability >= 0.75) |>
  group_by(day(date_time)) |>
  summarise(n = n())
```

#### Brown Creeper

```{r}
bc <- focused_time |> filter(species == "Certhia americana_Brown Creeper")

ggplot(data = bc, aes(x = time, y = id)) +
  geom_jitter(aes(colour = probability), alpha = 0.5, height = 0.25) +
  facet_wrap(~day(date_time))+
  theme_minimal() +
  scale_colour_viridis_c()
```

#### Rusty Blackbird

```{r}
rb <- focused_time |> filter(species == "Euphagus carolinus_Rusty Blackbird")

ggplot(data = rb, aes(x = time, y = id)) +
  geom_jitter(aes(colour = probability), alpha = 0.5, height = 0.25) +
  facet_wrap(~day(date_time))+
  theme_minimal() +
  scale_x_time(date_breaks = "30 min")+
  scale_colour_viridis_c()
```

#### Song sparrow

```{r}
sparrow <- focused_time |> filter(species == "Melospiza melodia_Song Sparrow")

ggplot(data = sparrow, aes(x = time, y = id)) +
  geom_jitter(aes(colour = probability), alpha = 0.5, height = 0.25) +
  facet_wrap(~day(date_time))+
  theme_minimal() +
  scale_colour_viridis_c()
```

#### Common Grackle

```{r}
grackle <- focused_time |> filter(species == "Quiscalus quiscula_Common Grackle")

ggplot(data = grackle, aes(x = time, y = id)) +
  geom_jitter(aes(colour = probability), alpha = 0.5, height = 0.25) +
  facet_wrap(~day(date_time))+
  theme_minimal() +
  scale_colour_viridis_c()
```


## Exploratory plots

Need to turn the data into binary data for logistic regression.

Currently the data has 'gaps' due to the times where nothing was detected being filtered out.

Another thing to not is that there are no seconds in any of the time or date-time variables, but I can add them using either the start or end time of each clip. I think the start time makes more sense.
```{r}
# eclipse time: 15:25:30
interval_start <- ymd_hms("2024-04-08 14:55:30")
interval_end <- ymd_hms("2024-04-08 15:55:30")

binary_full <- merged_df |> filter(date(date_time) == "2024-04-08") |> 
  filter(between(date_time, interval_start, interval_end)) |>
  complete(date_time = seq(interval_start, interval_end, by = "3 sec"), 
  fill = list(value = NA)) |> 
    mutate(Presence = if_else(!is.na(probability),
    true = 1,
    false = 0)) |> 
      relocate(Presence, probability) |> 
      filter(seconds(date_time) != 57)# we need to remove 57 sec pt from every minute


ggplot(binary_full, aes(as_hms(date_time), Presence))+
  geom_jitter(width = 0, height = 0.05)+
  geom_smooth()+
  theme_minimal()
```

Too hard to see. Let's look at some specific species:

#### Grackle

```{r}
binary_grackle <- merged_df |> filter(date(date_time) == "2024-04-08") |>
  filter(between(date_time, interval_start, interval_end)) |>
  filter(species == "Quiscalus quiscula_Common Grackle") |> 
  complete(date_time = seq(interval_start, interval_end, by = "3 sec"), 
  fill = list(value = NA)) |> 
    mutate(Presence = if_else(!is.na(probability),
    true = 1,
    false = 0)) |> 
      relocate(Presence, probability)

ggplot(binary_grackle, aes(as_hms(date_time), Presence))+
  geom_jitter(width = 0, height = 0.05)+
  geom_smooth()+
  theme_minimal()
```

#### Rusty Blackbird

```{r}
binary_rb <- merged_df |> filter(date(date_time) == "2024-04-08") |>
  filter(between(date_time, interval_start, interval_end)) |>
  filter(species == "Euphagus carolinus_Rusty Blackbird") |> 
  complete(date_time = seq(interval_start, interval_end, by = "3 sec"), 
  fill = list(value = NA)) |> 
    mutate(Presence = if_else(!is.na(probability),
    true = 1,
    false = 0)) |> 
      relocate(Presence, probability)

ggplot(binary_rb, aes(as_hms(date_time), Presence))+
  geom_jitter(width = 0, height = 0.05)+
  theme_minimal()
```

#### Song Sparrow

```{r}
binary_ss <- merged_df |> filter(date(date_time) == "2024-04-08") |>
  filter(between(date_time, interval_start, interval_end)) |>
  filter(species == "Melospiza melodia_Song Sparrow") |> 
  complete(date_time = seq(interval_start, interval_end, by = "3 sec"), 
  fill = list(value = NA)) |> 
    mutate(Presence = if_else(!is.na(probability),
    true = 1,
    false = 0)) |> 
      relocate(Presence, probability)

ggplot(binary_ss, aes(as_hms(date_time), Presence))+
  geom_jitter(width = 0, height = 0.05)+
  theme_minimal()
```

#### Brown Creeper

```{r}
binary_bc <- merged_df |> filter(date(date_time) == "2024-04-08") |>
  filter(between(date_time, interval_start, interval_end)) |>
  filter(species == "Certhia americana_Brown Creeper") |> 
  complete(date_time = seq(interval_start, interval_end, by = "3 sec"), 
  fill = list(value = NA)) |> 
    mutate(Presence = if_else(!is.na(probability),
    true = 1,
    false = 0)) |> 
      relocate(Presence, probability)

ggplot(binary_bc, aes(as_hms(date_time), Presence))+
  geom_jitter(width = 0, height = 0.05)+
  theme_minimal()
```

#### Robin

```{r}
binary_robin <- merged_df |> filter(date(date_time) == "2024-04-08") |>
  filter(between(date_time, interval_start, interval_end)) |>
  filter(species == "Turdus migratorius_American Robin") |> 
  complete(date_time = seq(interval_start, interval_end, by = "3 sec"), 
  fill = list(value = NA)) |> 
    mutate(Presence = if_else(!is.na(probability),
    true = 1,
    false = 0)) |> 
      relocate(Presence, probability)

ggplot(binary_robin, aes(as_hms(date_time), Presence))+
  geom_jitter(width = 0, height = 0.05)+
  theme_minimal()
```

#### Spring Peeper

```{r}
binary_sp <- merged_df |> filter(date(date_time) == "2024-04-08") |>
  filter(between(date_time, interval_start, interval_end)) |>
  filter(species == "Pseudacris crucifer_Spring Peeper") |> 
  complete(date_time = seq(interval_start, interval_end, by = "3 sec"), 
  fill = list(value = NA)) |> 
    mutate(Presence = if_else(!is.na(probability),
    true = 1,
    false = 0)) |> 
      relocate(Presence, probability)

ggplot(binary_sp, aes(as_hms(date_time), Presence))+
  geom_jitter(width = 0, height = 0.05)+
  geom_smooth()+
  theme_minimal()
```


## 5 day span

First, let's look at which species are the most common in the 5 surrounding days during the eclipse time:

```{r}
# 5 day span
day1 <- ymd_hms("2024-04-06 00:00:00")
day5 <- ymd_hms("2024-04-10 23:59:59")

# eclipse time: 15:25:30
time_start <- as_hms("14:55:30")
time_end <- as_hms("15:55:30")

five_day <- merged_df |> filter(between(time, 
time_start, time_end)) |>
  filter(day1 <= date_time & date_time <= day5) |>
  mutate(date = as_date(date_time)) |> 
  group_by(date, id) |> # want to complete the gaps for each site/AM
  complete(time = as_hms(seq(as.numeric(time_start), as.numeric(time_end), by = 3)), 
  fill = list(value = NA)) |> 
    ungroup() |> 
    mutate(Presence = if_else(!is.na(probability),
    true = 1,
    false = 0)) |> 
      relocate(Presence, probability) |> 
      filter(seconds(time) != 57)


# find the top 10 most common species
top_ten_5d <- five_day |> filter(!is.na(probability)) |> 
  group_by(species) |> 
  summarise(count = n()) |> 
  arrange(desc(count)) |> 
  slice(1:10)
top_ten_5d
```

Some new characters! Let's take a look at some of them and check if they have any missing audiomoths

## Missing AMs

There are 20 AMs in total

Let's look at the 10 ten species and see how many AMs are missing within the five days.

```{r}
five_day |> filter(species %in% top_ten_5d$species) |> 
  group_by(species) |>  
  summarise(missingAMs = 20 - n_distinct(id))
```

## Facet by surrounging days:

Let's make some plots of the surrounding days starting with all species, and then filtering to each of the native species in the 'top ten list.'

```{r}
ggplot(five_day, aes(time, Presence))+
  geom_jitter(width = 0, height = 0.05)+
  geom_smooth(se = F)+
  facet_wrap(~day(date))+
  scale_x_time()+
  theme_minimal()
```

#### Spring Peeper

```{r}
sp_5d <- five_day |> filter(species == "Pseudacris crucifer_Spring Peeper") |>
  group_by(date, id) |> 
  complete(time = as_hms(seq(as.numeric(time_start), as.numeric(time_end), by = 3)), 
  fill = list(value = NA)) |>
  mutate(Presence = if_else(!is.na(probability),
    true = 1,
    false = 0)) |> 
  filter(seconds(time) != 57)

ggplot(sp_5d, aes(time, Presence))+
  geom_jitter(width = 0, height = 0.05)+
  geom_smooth(se = F)+
  facet_wrap(~day(date))+
  scale_x_time()+
  theme_minimal()

# check which audiomoths had hits:
sp_5d |> group_by(id) |> 
  summarise(total_hits = sum(Presence))
```

Looks like we have increase activity around totality!

#### Wood Frog

```{r}
wf_5d <- five_day |> filter(species == "Lithobates sylvaticus_Wood Frog") |>
  group_by(date, id) |> 
  complete(time = as_hms(seq(as.numeric(time_start), as.numeric(time_end), by = 3)), 
  fill = list(value = NA)) |>
  mutate(Presence = if_else(!is.na(probability),
    true = 1,
    false = 0)) |> 
  filter(seconds(time) != 57)

ggplot(wf_5d, aes(time, Presence))+
  geom_jitter(width = 0, height = 0.05)+
  geom_smooth(se = F)+
  facet_wrap(~day(date))+
  scale_x_time()+
  theme_minimal()

# check which audiomoths had hits:
wf_5d |> group_by(id) |> 
  summarise(total_hits = sum(Presence))
```

No ids on the 6th (within the time interval). Hard to see a pattern due to the eclipse (possibly a drop in activity?)

#### Canada Goose

```{r}
cgo_5d <- five_day |> filter(species == "Branta canadensis_Canada Goose") |>
  group_by(date, id) |> 
  complete(time = as_hms(seq(as.numeric(time_start), as.numeric(time_end), by = 3)), 
  fill = list(value = NA)) |>
  mutate(Presence = if_else(!is.na(probability),
    true = 1,
    false = 0)) |> 
  filter(seconds(time) != 57)

ggplot(cgo_5d, aes(time, Presence))+
  geom_jitter(width = 0, height = 0.05)+
  geom_smooth(se = F)+
  facet_wrap(~day(date))+
  scale_x_time()+
  theme_minimal()

# check which audiomoths had hits:
cgo_5d |> group_by(id) |> 
  summarise(total_hits = sum(Presence))
```

Seems like there is a gap in activity around totality that is not present in the other days.

#### Pickerel Frog

```{r}
pf_5d <- five_day |> filter(species == "Lithobates palustris_Pickerel Frog") |>
  group_by(date, id) |> 
  complete(time = as_hms(seq(as.numeric(time_start), as.numeric(time_end), by = 3)), 
  fill = list(value = NA)) |>
  mutate(Presence = if_else(!is.na(probability),
    true = 1,
    false = 0)) |> 
  filter(seconds(time) != 57)

ggplot(pf_5d, aes(time, Presence))+
  geom_jitter(width = 0, height = 0.05)+
  geom_smooth(se = F)+
  facet_wrap(~day(date))+
  scale_x_time()+
  theme_minimal()

# check which audiomoths had hits:
cgo_5d |> group_by(id) |> 
  summarise(total_hits = sum(Presence))
```

No ids on the 6th (within the time interval). Maybe a little increase in activity around totality but looks minimal.

#### Chickadee

```{r}
bcc_5d <- five_day |> filter(species == "Poecile atricapillus_Black-capped Chickadee") |>
  group_by(date, id) |> 
  complete(time = as_hms(seq(as.numeric(time_start), as.numeric(time_end), by = 3)), 
  fill = list(value = NA)) |>
  mutate(Presence = if_else(!is.na(probability),
    true = 1,
    false = 0)) |> 
  filter(seconds(time) != 57)

ggplot(bcc_5d, aes(time, Presence))+
  geom_jitter(width = 0, height = 0.05)+
  geom_smooth(se = F)+
  facet_wrap(~day(date))+
  scale_x_time()+
  theme_minimal()

# check which audiomoths had hits:
bcc_5d |> group_by(id) |> 
  summarise(total_hits = sum(Presence))
```

Data is pretty sparse on the 7th and 8th -- probably a difficult species to analyze for these dates within the time interval.

#### Common Grackle

```{r}
cgr_5d <- five_day |> filter(species == "Quiscalus quiscula_Common Grackle") |>
  group_by(date, id) |> 
  complete(time = as_hms(seq(as.numeric(time_start), as.numeric(time_end), by = 3)), 
  fill = list(value = NA)) |>
  mutate(Presence = if_else(!is.na(probability),
    true = 1,
    false = 0)) |> 
  filter(seconds(time) != 57)

ggplot(cgr_5d, aes(time, Presence))+
  geom_jitter(width = 0, height = 0.05)+
  geom_smooth(se = F)+
  facet_wrap(~day(date))+
  scale_x_time()+
  theme_minimal()

# check which audiomoths had hits:
cgr_5d |> group_by(id) |> 
  summarise(total_hits = sum(Presence))
```

Data pretty sparse for the time interval on the 6th. There seems to be a gap around totality on the 8th, which is not present for the other days.


Palmeria dolei_Akohekohe:

```{r}
a_5d <- five_day |> filter(species == "Palmeria dolei_Akohekohe") |>
  group_by(date, id) |> 
  complete(time = as_hms(seq(as.numeric(time_start), as.numeric(time_end), by = 3)), 
  fill = list(value = NA)) |>
  mutate(Presence = if_else(!is.na(probability),
    true = 1,
    false = 0)) |> 
  filter(seconds(time) != 57)

ggplot(a_5d, aes(time, Presence))+
  geom_jitter(width = 0, height = 0.05)+
  geom_smooth(se = F)+
  facet_wrap(~day(date))+
  scale_x_time()+
  theme_minimal()

# check which audiomoths had hits:
a_5d |> group_by(id) |> 
  summarise(total_hits = sum(Presence))
```

Porzana porzana_Spotted Crake

```{r}
sc_5d <- five_day |> filter(species == "Porzana porzana_Spotted Crake") |>
  group_by(date, id) |> 
  complete(time = as_hms(seq(as.numeric(time_start), as.numeric(time_end), by = 3)), 
  fill = list(value = NA)) |>
  mutate(Presence = if_else(!is.na(probability),
    true = 1,
    false = 0)) |> 
  filter(seconds(time) != 57)

ggplot(sc_5d, aes(time, Presence))+
  geom_jitter(width = 0, height = 0.05)+
  geom_smooth(se = F)+
  facet_wrap(~day(date))+
  scale_x_time()+
  theme_minimal()

# check which audiomoths had hits:
sc_5d |> group_by(id) |> 
  summarise(total_hits = sum(Presence))
```


