---
title: "Modeling"
format: html
---

## Set up

```{r}
#| warning: false
#| output: false
rm(list = ls())
library(tidyverse)
library(here)
library(nlme)
library(mgcv)
library(forecast)
library(hms) 
merged_df <- read_csv(here("Data/merged_output_deployment.csv"))
```

#### Fix datetime variable

```{r}
merged_df <- merged_df |> 
    mutate(date_time = date_time + seconds(start),
    time = as_hms(time + start))
```

## Single AM on Eclipse Day

Here we will do a simplified model where we will only look at one site on eclipse day but look at all species.

First we need to filter the data frame and fill in the gaps in time where there were no species calling.

#### Filtering Data Frame 

We will use deployment AM0011

```{r}
# eclipse time: 15:25:30
time_start <- as_hms("14:55:30")
time_end <- as_hms("15:55:30")

AM0011_all_sp <- merged_df |> filter(as_date(date_time) == "2024-04-08" & deployment_num == "AM0011") |>   
    filter(between(time, time_start, time_end)) |>
  complete(time = as_hms(seq(as.numeric(time_start), as.numeric(time_end), by = 3)), 
  fill = list(value = NA)) |> 
    ungroup() |> 
    mutate(Presence = if_else(!is.na(probability),
    true = 1,
    false = 0)) |> 
      relocate(Presence, probability) |> 
      filter(seconds(time) != 57)
```

Looks good! Now we can move on to modelling!

### Autocorrelation

```{r}
ts_AM0011 <- ts(AM0011_all_sp$Presence, start = 53730, end = 57330, deltat = 3)
Acf(ts_AM0011)
Pacf(ts_AM0011)
```

Definitely need at least on AR term.

### Model

```{r}
# some timestamps have multiple species (need to get rid of duplicates)
AM0011_all_sp <- AM0011_all_sp |> distinct(time, .keep_all = TRUE)

# need time variable to be numeric:
AM0011_all_sp$time_num <- as.numeric(AM0011_all_sp$time)

AM0011_mod <- gamm(Presence ~ s(time_num), data = AM0011_all_sp, family = binomial, correlation = corAR1(form = ~ time_num))
summary(AM0011_mod$gam)
plot(AM0011_mod$gam)
```

#### Autocorrelation of resids

```{r}
resids_AM0011 <- residuals(AM0011_mod$lme)
Acf(resids_AM0011)
```


## All AMs on Eclipse Day

Next we will use the same model to look at all species from all the sites, but still only on eclipse day.

### Filter DF

```{r}
# eclipse time: 15:25:30
time_start <- as_hms("14:55:30")
time_end <- as_hms("15:55:30")

all_sites <- merged_df |> filter(as_date(date_time) == "2024-04-08") |> 
    filter(between(time, time_start, time_end)) |>
    group_by(deployment_num) |> # want to complete the gaps for each site/AM
    complete(time = as_hms(seq(as.numeric(time_start), as.numeric(time_end), by = 3)), 
    fill = list(value = NA)) |> 
    ungroup() |> 
    mutate(Presence = if_else(!is.na(probability),
    true = 1,
    false = 0)) |> 
      relocate(Presence, probability) |> 
      filter(seconds(time) != 57) |> 
    distinct(time, .keep_all = TRUE) # get rid of any species at the same timestamp
```

### Model

```{r}
# 1. Ensure site is a factor and data is sorted
all_sites$deployment_num <- as.factor(all_sites$deployment_num)

# 2. Create the within-site observation order to avoid duplicate timestamp errors
all_sites <- all_sites |> 
  group_by(deployment_num) |> 
  mutate(obs_order = row_number()) |> 
  ungroup()

all_sites$time_num <- as.numeric(all_sites$time)

# 3. Fit the model
all_sites_mod <- gamm(Presence ~ s(time_num) + s(deployment_num, bs = "re"), # re or sz?s
                   family = binomial,
                   data = all_sites,
                   correlation = corAR1(form = ~ obs_order | deployment_num))
summary(all_sites_mod$gam)
plot(all_sites_mod$gam)
```

Both models seem to be working?
